Cohesiveness in microbial community coalescence

### 2021-05-18 draft/draft_v2.tex
playing around with python code, checks to quantify bottom-up and top-down cohesiveness

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{makecell} % more table customization
\usepackage{mathtools} % nice matrices
\usepackage{units} % nice fractions
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\zxrsetup{toltxlabel}
\zexternaldocument*{supp}
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage[table,xcdraw]{xcolor} % text and table background colors
\usepackage[margin=3cm]{geometry} % margins
\usepackage{lineno} % line numbers
\usepackage{textcomp} % special characters (some references break if this isn't loaded?)
\usepackage{mdframed} % to easily handle boxes

@@

% title, authors, affiliations
\title{Cohesiveness in microbial community coalescence}
\author[1,*]{Juan Diaz-Colunga}
\author[1,*]{Nanxi Lu}
\author[1,2,*]{Alicia Sanchez-Gorostiaga}
\author[3]{Joshua E. Goldford}
\author[4]{Mikhail Tikhonov}
\author[1,$\dagger$]{Álvaro Sánchez}
\affil[1]{Department of Ecology \& Evolutionary Biology
and
Microbial Sciences Institute,
Yale University, New Haven, CT, USA}
\affil[2]{Department of Microbial Biotechnology,
Centro Nacional de Biotecnología (CNB-CSIC), Cantoblanco, Madrid, Spain}
\affil[3]{Physics of Living Systems, Department of Physics,
Massachusetts Institute of Technology, Cambridge, MA, USA}
\affil[4]{Department of Physics,
Center for Science \& Engineering of Living Systems,
Washington University in St. Louis, St. Louis, MO, USA}

\affil[$\dagger$]{\normalfont To whom correspondence should be addressed: alvaro.sanchez@yale.edu}
\affil[*]{\normalfont These authors contributed equally}
\date{}

@@

Our results indicate that co-selection varies in direction and strength depending on the supplied limiting
resource.
This primary resource, in turn, has been shown to
shape the structure and interactions of the communities \cite{Estrela2020}.
We observe that, when top-down co-selection is weak, bottom-up co-selection can be very strong,
with positive co-selection being far more common than negative co-selection.
\iffalse
Our results confirm that positive top-down co-selection is common, but its effects are weak.
In contrast, bottom-up co-selection can be very strong, and positive co-selection is far more common
than negative co-selection.
\fi
We then turn to a Microbial Consumer-Resource Model (MicroCRM)
\cite{Goldford2018,Marsland2019,Marsland2020}
that is able to capture the dynamics of microbial communities dominated by metabolic interactions,
as is the case for the ones assembled in our experimental conditions \cite{Goldford2018,Estrela2020}.
We show that the empirically observed trends in ecological co-selection are reproduced
with minimal model assumptions, and that
tuning the complexity of the metabolic interactions in our \textit{in silico} communities can modulate
the recurrence of top-down or bottom-up co-selection.
Our findings indicate that collective interactions play an important role at dictating community structure
during coalescence.

@@

\bigskip
\begin{center}
\includegraphics[width=8cm,keepaspectratio]{figs/figBox_v2.png}
\end{center}
\bigskip

>>>>

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{makecell} % more table customization
\usepackage{mathtools} % nice matrices
\usepackage{units} % nice fractions
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\zxrsetup{toltxlabel}
\zexternaldocument*{supp}
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage[table,xcdraw]{xcolor} % text and table background colors
\usepackage[margin=2.5cm]{geometry} % margins
\usepackage{lineno} % line numbers
\usepackage{textcomp} % special characters (some references break if this isn't loaded?)
\usepackage{mdframed} % to easily handle boxes
\usepackage[misc]{ifsym} % use \Letter symbol for corresponding author

@@

% title, authors, affiliations
\title{Cohesiveness in microbial community coalescence}
\author[1,*]{Juan Diaz-Colunga}
\author[1,*]{Nanxi Lu}
\author[1,2,*]{Alicia Sanchez-Gorostiaga}
\author[1]{Chang-Yu Chang}
\author[1]{Helen S. Cai}
\author[3]{Joshua E. Goldford}
\author[4]{Mikhail Tikhonov}
\author[1,\Letter]{Álvaro Sánchez}
\affil[1]{Department of Ecology \& Evolutionary Biology
and
Microbial Sciences Institute,
Yale University, New Haven, CT, USA}
\affil[2]{Department of Microbial Biotechnology,
Centro Nacional de Biotecnología (CNB-CSIC), Cantoblanco, Madrid, Spain}
\affil[3]{Physics of Living Systems, Department of Physics,
Massachusetts Institute of Technology, Cambridge, MA, USA}
\affil[4]{Department of Physics,
Center for Science \& Engineering of Living Systems,
Washington University in St. Louis, St. Louis, MO, USA}

\affil[$\textrm{\Letter}$]{\normalfont To whom correspondence should be addressed: alvaro.sanchez@yale.edu}
\affil[*]{\normalfont These authors contributed equally}
\date{}

@@

Our results indicate that co-selection varies in direction and strength depending on the supplied limiting
resource.
This primary resource, in turn, has been shown to
shape the structure and interactions of the communities \cite{Estrela2020}.
We observe that, when top-down co-selection is weak, bottom-up co-selection can be very strong,
with positive co-selection being far more common than negative co-selection.
\iffalse
Our results confirm that positive top-down co-selection is common, but its effects are weak.
In contrast, bottom-up co-selection can be very strong, and positive co-selection is far more common
than negative co-selection.
\fi
We then turn to a Microbial Consumer-Resource Model (MicroCRM)
\cite{Goldford2018,Marsland2019,Marsland2020}
that is able to capture the dynamics of microbial communities dominated by metabolic interactions,
as is the case for the ones assembled in our experimental conditions \cite{Goldford2018,Estrela2020}.
We show that the empirically observed trends in ecological co-selection are reproduced
with minimal model assumptions, and that the recurrence of top-down and bottom-up co-selection
is determined by the configuration of the cross-feeding networks in the MicroCRM.
Our findings indicate that collective interactions play an important role at dictating community structure
during coalescence.

@@

\bigskip
\begin{center}
\includegraphics[width=11cm,keepaspectratio]{figs/figBox_v3.png}
\end{center}
\bigskip
